Kronos \& Krios: Evidence for accretion of a massive, rocky planetary system in a comoving pair of solar-type stars

### 2017-09-11 paper/ms.tex
Minor edits to clarify; corner attribution

\newcommand{\todo}[1]{{\color{blue}TODO: #1\\}}
\newcommand{\apw}[1]{{\color{purple}APW asks: #1}}
\newcommand*{\askjmb}[1]{{\bf #1}}

@@

\begin{abstract}
  We report and discuss the discovery of a comoving pair of bright solar-type
  stars, HD~240430 and HD~240429, with a significant difference in their
  chemical abundances.
  The two stars have an estimated 3D separation of $\approx 0.6$~pc ($\approx 0.01$~pc tangentially) at a distance of
  $r\approx 100~\pc$ with nearly identical three-dimensional velocities,
  %($|\Delta \bs{v}| < XX~\kms$ with 95\% confidence),
  as inferred from \gaia\ \tgas\ parallaxes and proper motions, and
  high-precision radial velocity measurements.
  Stellar parameters determined from high-resolution Keck HIRES spectra
  indicate that both stars are $\sim 4$~Gyr old.
  The more metal-rich of the two, HD~240430, shows an enhancement of refractory
  ($\Tcondens>1200$~K) elements by $\approx 0.2$~dex and a marginal
  enhancement of (moderately) volatile elements ($\Tcondens<1200$~K; \elem{C},
  \elem{N}, \elem{O}, \elem{Na}, and \elem{Mn}).
  % TODO: Hmmm...
  This is the largest and the most significant chemical difference found in
  a wide binary pair yet.
  Additionally, HD~240430 shows an anomalously high surface lithium abundance
  ($A(\elem{Li})=2.75$), higher by $0.5$~dex than its companion.
  The proximity in phase-space and ages between the two stars suggests that
  they formed together with the same composition, at odds with the observed
  differences in metallicity and abundance patterns.
  We therefore suggest that the star HD~240430, ``Kronos'', accreted
  \maccreted\ of rocky material after birth, selectively enhancing the
  refractory elements as well as lithium in its surface and convective
  envelope.
\end{abstract}

@@

High resolution spectroscopic studies of binary star systems hosting at least
one planet (\citealt{2011ApJ...740...76R,2014ApJ...790L..25T,Teske:2013aa,
  Mack:2014aa,Liu:2014aa,Teske:2015aa,Saffe:2015aa,
  Ramirez:2015aa,Biazzo:2015aa,Mack:2016aa,Teske:2016aa,Teske:2016ab})
have yielded varied results:
while some systems appear to have undetectable differences
(see also \citealt{Desidera:2004aa,Gratton:2001aa}), other
studies have reported a \Tcondens-dependent difference in abundance
with higher-\Tcondens\ elements showing larger differences.
A possible explanation for the difference is that
forming more gas giants or rocky planets leads to
an overall or \Tcondens-dependent depletion of higher-\Tcondens\ elements that
eventually accrete onto the host star
(\citealt{Ramirez:2015aa,Biazzo:2015aa}).
Alternatively, late time accretion of refractory-rich planetary material to the
relatively refractory-rich star can also produce the trend \apw{do you mean "relatively refractory-poor star" instead?}.
The observed differences are $\lesssim 0.1$~dex even in the most dramatic case,
and often closer to $\approx 0.05$~dex, making them challenging to detect even
with a careful analysis of high-resolution, high signal-to-noise ratio spectra,
and differential analyses of two stars that are very similar in their
stellar parameters.
We refer the readers to Appendix~\ref{app:review} for a review of
a handful of individual pairs studied in their detailed chemical abundances
(see also \citealt{2016arXiv161104064M}).

@@

%TODO: corner.py
%TODO: data/code available
\software{
  %The data and code used in this project is available from
  %\url{https://github.com/smoh/KronosKrios} under the MIT open-source
  %software license.
  This research utilized:
  \texttt{Astropy} (\citealt{Astropy-Collaboration:2013}),
  \texttt{emcee} (\citealt{2013PASP..125..306F}),
  \texttt{IPython} (\citealt{Perez:2007}),
  \texttt{matplotlib} (\citealt{Hunter:2007}),
  \texttt{numpy} (\citealt{Van-der-Walt:2011}),
  and \texttt{pandas} (\citealt{pandas}).}

>>>>

\newcommand{\todo}[1]{{\color{blue}TODO: #1\\}}
\newcommand*{\askjmb}[1]{{\bf #1}}

@@

\begin{abstract}
  We report and discuss the discovery of a comoving pair of bright solar-type
  stars, HD~240430 and HD~240429, with a significant difference in their
  chemical abundances.
  The two stars have an estimated 3D separation of $\approx 0.6$~pc
  ($\approx 0.01$~pc projected) at a distance of
  $r\approx 100~\pc$ with nearly identical three-dimensional velocities,
  %($|\Delta \bs{v}| < XX~\kms$ with 95\% confidence),
  as inferred from \gaia\ \tgas\ parallaxes and proper motions, and
  high-precision radial velocity measurements.
  Stellar parameters determined from high-resolution Keck HIRES spectra
  indicate that both stars are $\sim 4$~Gyr old.
  The more metal-rich of the two, HD~240430, shows an enhancement of refractory
  ($\Tcondens>1200$~K) elements by $\approx 0.2$~dex and a marginal
  enhancement of (moderately) volatile elements ($\Tcondens<1200$~K; \elem{C},
  \elem{N}, \elem{O}, \elem{Na}, and \elem{Mn}).
  % TODO: Hmmm...
  This is the largest and the most significant chemical difference found in
  a wide binary pair yet.
  Additionally, HD~240430 shows an anomalously high surface lithium abundance
  ($A(\elem{Li})=2.75$), higher by $0.5$~dex than its companion.
  The proximity in phase-space and ages between the two stars suggests that
  they formed together with the same composition, at odds with the observed
  differences in metallicity and abundance patterns.
  We therefore suggest that the star HD~240430, ``Kronos'', accreted
  \maccreted\ of rocky material after birth, selectively enhancing the
  refractory elements as well as lithium in its surface and convective
  envelope.
\end{abstract}

@@

High resolution spectroscopic studies of binary star systems hosting at least
one planet (\citealt{2011ApJ...740...76R,2014ApJ...790L..25T,Teske:2013aa,
  Mack:2014aa,Liu:2014aa,Teske:2015aa,Saffe:2015aa,
  Ramirez:2015aa,Biazzo:2015aa,Mack:2016aa,Teske:2016aa,Teske:2016ab})
have yielded varied results:
while some systems appear to have undetectable differences
(see also \citealt{Desidera:2004aa,Gratton:2001aa}), other
studies have reported a \Tcondens-dependent difference in abundance
with higher-\Tcondens\ elements showing larger differences.
A possible explanation for the difference is that
forming more gas giants or rocky planets leads to
an overall or \Tcondens-dependent depletion of high-\Tcondens\ elements in
gas that eventually accretes onto the host star
(\citealt{Ramirez:2015aa,Biazzo:2015aa}).
Alternatively, late time accretion of refractory-rich planetary material can
also produce the trend by enhancing the abundance of high-\Tcondens\ elements
in one of the two stars instead.
The observed differences are $\lesssim 0.1$~dex even in the most dramatic case,
and often closer to $\approx 0.05$~dex, making them challenging to detect even
with a careful analysis of high-resolution, high signal-to-noise ratio spectra,
and differential analyses of two stars that are very similar in their
stellar parameters.
We refer the readers to Appendix~\ref{app:review} for a review of
a handful of individual pairs studied in their detailed chemical abundances
(see also \citealt{2016arXiv161104064M}).

@@

%TODO: data/code available
\software{
  %The data and code used in this project is available from
  %\url{https://github.com/smoh/KronosKrios} under the MIT open-source
  %software license.
  This research utilized:
  \texttt{Astropy} (\citealt{Astropy-Collaboration:2013}),
  \texttt{corner.py} (\citealt{corner}),
  \texttt{emcee} (\citealt{2013PASP..125..306F}),
  \texttt{IPython} (\citealt{Perez:2007}),
  \texttt{matplotlib} (\citealt{Hunter:2007}),
  \texttt{numpy} (\citealt{Van-der-Walt:2011}),
  and \texttt{pandas} (\citealt{pandas}).}
